R functional programming approach for truncated and shifted and for shifted and truncated probability distribution functions

### 2016-09-21 docs/shift+trunc-article/article-shift+trunc.Rmd
s

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

R programming functional nature allows programmer to implement and manipulate functions as needed. Functional programming is a style of programming that emphasizes the evaluation of expressions rather than the execution of commands, so these functions produce results that depend only on their inputs and not on the program state. This style has a great power and includes lots of tools that combined together can allows us to produce results with the most efficient computational method. One of them is the functional factory, which is a factory for making new functions. It allows us to use closures, which are functions together with a referencing environment, to write specific functions that, in turn, can be used to generate new functions. So we can have a double layer of development: a first layer that is used to do all the complex work in common to all functions and a second layer that defines the details of each function. This fundamentals was used to develop the computational method proposed by Andrea Spanò in "R package qdist: A functional programming approach to truncated probability distributions", and that I used in "R package qshift: A functional programming approach to shifted probability distributions", that produces functions for truncated and shifted distributions, respectively.  [`qdist`](https://github.com/veronicagiro/qdist) and [`qshift`](https://github.com/veronicagiro/qshift) packages are available on Github:

```{r, eval=FALSE}
devtools::install_github("veronicagiro/qtruncate")
devtools::install_github("veronicagiro/qshift")
```

In the previous citated articles, we dealt separately with shifted and truncated distributions. But in real life, we sometimes work with random variables arising from more than one transformation. So in this article, we will demonstrate that we can produce functions for truncated and shifted or for shifted and truncated distributions only combining the functions of the two packages. This is possible because of functional programming approach used to build them.

@@

* `d<dist>(x)` with `x` vector of quantiles, identifies density functions
* `p<dist>(q)` with `q` vector of quantiles, identifies probability distribution functions
* `q<dist>(p)` with `p` vector of probabilities, identifies quantile functions
* `r<dist>(n)` with `n` number of observations, identifies random number generation

@@

Andrea Spanò proposed in `qdist` a functional programming method that computes truncated distributions functions for each distribution chosen without changing the well structured programming style and syntax of previous functions.

@@

* the environment of a function can be used as a placeholder for other objects (for detail see Andrea Spanò article "R package qdist: A functional programming approach to truncated probability ditributions"")
* function formals can be manipulated

@@

1. defines the required functions suitable for computing from specific truncated probability distribution by using the functions available within its package. These functions take as input a probability distribution as a character string, in this case the Weibull distribution, and return the equivalent truncated distribution function as object:

```{r, message=FALSE}
require(qdist)
dtweibull <- dtruncate("weibull") # density function
ptweibull <- ptruncate("weibull") # probability function 
qtweibull <- qtruncate("weibull") # quantile function
rtweibull <- rtruncate("weibull") # random generator function
```

@@

```{r}
args(dweibull)
args(dtweibull)
```

As `dweibull()` is the function that computes the density function for a Weibull distribution, conventionally named `weibull`, `dtweibull()` is the function that computes the density function for a truncated Weibull distributions, named `tweibull`.

Suppose we want to generate the density function for a **truncated and shifted** Weibull distribution. We need `qshift` package. `qshift` package shares the same logic and the same methodology of `qdist`. The difference is that the functions are designed to compute functions for shifted distributions.
In particular, `qshift` package includes the following four functions:

@@

```{r, message=FALSE}
require(qshift)
dstweibull <- dshift("tweibull")
```

@@

```{r}
args(dstweibull)
```

@@

```{r}
x <- seq(0, 5, len = 100)
orig_density <- dweibull(x = x, shape = 1.5, scale = 1)
trunc_density <- dtweibull(x = x, shape = 1.5, scale = 1, L = 1.5, U = 3)
trunc_shift_density <- dstweibull(x = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r, message=FALSE}
df <- data.frame(x=x, 
                 type = factor(c(rep("orig_density", times=length(x)),
                    rep("trunc_density", times=length(x)),
                    rep("trunc_shift_density", times=length(x))),
                    levels = c("orig_density", "trunc_density", "trunc_shift_density")), 
                  value = c(orig_density, trunc_density, trunc_shift_density))
require(ggplot2)
pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("orig_density" = "cyan3", 
                                "trunc_density" = "#89FF99", 
                                "trunc_shift_density" = "#FF89B4")) +
  xlab("Value") + ylab("Density") + ylim(c(0,1.9)) +
  ggtitle(paste("Density plot for non-transformed, truncated", "and truncated and shifted Weibull distributions", sep = "\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

@@

Starting from `dshift()` function of `qshift` package, we define `dsweibull()`, the density function of `sweibull`, which is the shifted Weibull distribution:
```{r}
dsweibull <- dshift("weibull") # density function
```

We have to define also the others functions for shifted distribution, as they are implicitly necessary for the further step:

```{r}
psweibull <- pshift("weibull") # probability function 
qsweibull <- qshift("weibull") # quantile function
rsweibull <- rshift("weibull") # random generator function
```

The density function of a shifted and truncated Weibull distribution is generated setting `sweibull` distribution as `ptruncate()` argument:
```{r}
dtsweibull <- dtruncate("sweibull")
```

@@

```{r}
orig_density <- dweibull(x = x, shape = 1.5, scale = 1)
shift_density <- dsweibull(x = x, shape = 1.5, scale = 1, shift = 1)
shift_trunc_density <- dtsweibull(x = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r}
df <- data.frame(x=x, 
                 type = factor(c(rep("orig_density", times=length(x)),
                    rep("shift_density", times=length(x)),
                    rep("shift_trunc_density", times=length(x))),
                    levels = c("orig_density", "shift_density", "shift_trunc_density")), 
                  value = c(orig_density, shift_density, shift_trunc_density))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("orig_density" = "cyan3", 
                                "shift_density" = "#89FF99", 
                                "shift_trunc_density" = "#FF89B4")) +
  xlab("Value") + ylab("Density") + ylim(c(0,1.9)) +
  ggtitle(paste("Density plot for non-transformed, shifted and","shifted and truncated Normal distributions", sep = "\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

@@

Let us see graphically the differences between the density function of Weibull distribution, of a truncated and shifted Weibull distribution and of a shifted and truncated Weibull distribution.

```{r}
df <- data.frame(x=x, 
                 type = factor(c(rep("orig_density", times=length(x)),
                    rep("trunc_shift_density", times=length(x)),
                    rep("shift_trunc_density", times=length(x))),
                    levels = c("orig_density", "trunc_shift_density", "shift_trunc_density")), 
                  value = c(orig_density, trunc_shift_density, shift_trunc_density))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("orig_density" = "cyan3", 
                                "trunc_shift_density" = "#89FF99", 
                                "shift_trunc_density" = "#FF89B4")) +
  xlab("Value") + ylab("Density") + ylim(c(0,1.9)) +
  ggtitle(paste("Density plot for non-transformed, truncated-shifted", 
                "and shifted-truncated Weibull distribution", sep="\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

@@

Looking at the figure we can see that a truncated and shifted distribution is different from a shifted amd truncated distribution.

Let us analyze the other distributions functions:

@@

```{r}
pstweibull <- pshift("tweibull")
ptsweibull <- ptruncate("sweibull")
orig_distr <- pweibull(q = x, shape = 1.5, scale = 1)
shift_trunc_distr <- pstweibull(q = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
trunc_shift_distr <- ptsweibull(q = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r}
df <- data.frame(x=x, 
                 type = factor(c(rep("orig_distr", times=length(x)),
                    rep("shift_trunc_distr", times=length(x)),
                    rep("trunc_shift_distr", times=length(x))),
                    levels = c("orig_distr", "shift_trunc_distr", "trunc_shift_distr")), 
                  value = c(orig_distr, shift_trunc_distr, trunc_shift_distr))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("orig_distr" = "cyan3", 
                                "shift_trunc_distr" = "#89FF99", 
                                "trunc_shift_distr" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle(paste("Distribution plot for non-transformed, truncated-shifted", 
                "and shifted-truncated Weibull distribution", sep="\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

@@

```{r}
qstweibull <- qshift("tweibull")
qtsweibull <- qtruncate("sweibull")
x <- 1:999/1000
orig_quant <- qweibull(p = x, shape = 1.5, scale = 1)
shift_trunc_quant <- qstweibull(p = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
trunc_shift_quant <- qtsweibull(p = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r}
df <- data.frame(x=x, 
                 type = factor(c(rep("orig_quant", times=length(x)),
                    rep("shift_trunc_quant", times=length(x)),
                    rep("trunc_shift_quant", times=length(x))),
                    levels = c("orig_quant", "shift_trunc_quant", "trunc_shift_quant")), 
                  value = c(orig_quant, shift_trunc_quant, trunc_shift_quant))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("orig_quant" = "cyan3", 
                                "shift_trunc_quant" = "#89FF99", 
                                "trunc_shift_quant" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle(paste("Quantile plot for non-transformed, truncated-shifted", 
                "and shifted-truncated Weibull distribution", sep="\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

@@

```{r}
rstweibull <- rshift("tweibull")
rtsweibull <- rtruncate("sweibull")
orig_rand <- rweibull(n=1000, shape = 1.5, scale = 1)
shift_trunc_rand <- rstweibull(n=1000, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
trunc_shift_rand <- rtsweibull(n=1000, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r}
df <- data.frame(x=1:1000, 
                 type = factor(c(rep("orig_rand", times=1000),
                    rep("shift_trunc_rand", times=1000),
                    rep("trunc_shift_rand", times=1000)),
                    levels = c("orig_rand", "shift_trunc_rand", "trunc_shift_rand")), 
                  value = c(orig_rand, shift_trunc_rand, trunc_shift_rand))

pl <- ggplot(data = df, mapping = aes(x=value, group=type, fill=type)) +
        geom_density(alpha=0.6) +
        scale_fill_manual(values = c("orig_rand" = "cyan3",
                              "shift_trunc_rand" = "#89FF99",
                              "trunc_shift_rand" = "#FF89B4")) +
        ggtitle(paste("Density plot for non-transformed, truncated-shifted", 
                      "and shifted-truncated Weibull distribution", sep="\n")) +
        theme(legend.title=element_blank(), legend.text= element_text(size=10),
        plot.title = element_text(size=14), axis.title=element_text(size=12))

@@

```{r}
ltweibull <- function(x, shift = 0 ,L = -Inf, U = Inf){
  x1 <- x - shift
  # starting values for parameters (scale, shape)
  shape <- (sd(x1)/mean(x1))^(-1.086)
  scale <- mean(x1)/gamma(1+1/shape)
  # parameters vector definition
  theta <- c(shape, scale)
  # likelihood function
  ll <- function(theta, x, shift = 0, L = -Inf, U = Inf){
    shape <- theta[1]
    scale <- theta[2]
    ld <- dstweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U, log=TRUE)
    -sum(ld)
  }
  # maximum likelihood estimation
  optim(par = theta, fn = ll, x = x, shift=shift, L = L, U = U, method = "Nelder-Mead")[["par"]]
}
```

@@

```{r}
x <- rstweibull(n = 100000, shape = 1.6, scale = 1, shift = 1, L = 1, U = 2.5)
# maximum likelihood estimate for the shifted and truncated weibull distribution
ltweibull(x = x, shift=1, L = 1, U = 2.5)
```

@@

We will consider the previously defined density function, `dtsweibull()`, for a shifted and truncated Weibull distribution. We use `dtsweibull()` inside the estimator function:

```{r}
ltweibull <- function(x, shift = 0, L = -Inf, U = Inf){
  x1 <- x - shift
  # starting values for parameters (scale, shape)
  shape <- (sd(x1)/mean(x1))^(-1.086)
  scale <- mean(x1)/gamma(1+1/shape)
  # parameters vector definition
  theta <- c(shape, scale)
  # likelihood function
  ll <- function(theta, x, L = -Inf, U = Inf){
    shape <- theta[1]
    scale <- theta[2]
    ld <- dtsweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U, log=T)
    -sum(ld)
  }
  # maximum likelihood estimation
  optim(par = theta, fn = ll, x = x, L = L, U = U, method = "Nelder-Mead")[["par"]]
}
```

@@

```{r}
x <- rtsweibull(n = 100000, shape = 1.6, scale = 1, shift = 1, L = 0.5, U = 3)
# maximum likelihood estimate for the shifted weibull distribution
ltweibull(x = x, shift = 1, L = 0.5, U = 3)
```

>>>>

@@

R programming functional nature allows programmer to implement and manipulate functions as needed. Functional programming is a style of programming that emphasizes the evaluation of expressions rather than the execution of commands, so these functions produce results that depend only on their inputs and not on the program state. This style has a great power and includes lots of tools that combined together can allows us to produce results with the most efficient computational method. One of them is the functional factory, which is a factory for making new functions. It allows us to use closures, which are functions together with a referencing environment, to write specific functions that, in turn, can be used to generate new functions. So we can have a double layer of development: a first layer that is used to do all the complex work in common to all functions and a second layer that defines the details of each function. This concept was used to develop the computational method proposed by Andrea Spanò in "R package qdist: A functional programming approach to truncated probability distributions", and that I used in "R package qshift: A functional programming approach to shifted probability distributions", that produces functions for truncated and shifted distributions, respectively.  [`qdist`](https://github.com/veronicagiro/qdist) and [`qshift`](https://github.com/veronicagiro/qshift) packages are available on Github:

```{r repos, eval=FALSE}
devtools::install_github("veronicagiro/qtruncate")
devtools::install_github("veronicagiro/qshift")
```

In the previously cited articles, we dealt separately with shifted and truncated distributions. But in real life, we sometimes work with random variables arising from more than one transformation. So in this article, we will demonstrate that we can produce functions for truncated and shifted or for shifted and truncated distributions only combining the functions of the `qdist` and `qshift` packages. This is possible because of functional programming approach used to build them.

@@

* `d<dist>(x)` with `x` vector of quantiles, identifies density functions
* `p<dist>(q)` with `q` vector of quantiles, identifies probability distribution functions
* `q<dist>(p)` with `p` vector of probabilities, identifies quantile functions
* `r<dist>(n)` with `n` number of observations, identifies random number generator functions

@@

Andrea Spanò proposed in `qdist` a functional programming method that computes truncated distribution functions for each distribution chosen without changing the well structured programming style and syntax of the previous functions.

@@

* the environment of a function can be used as a placeholder for other objects (for details see Andrea Spanò article "R package qdist: A functional programming approach to truncated probability distributions")
* function formals can be manipulated

@@

1. defines the required functions suitable for computing from a specific truncated probability distribution by using the functions available within its package. These functions take as input a probability distribution as a character string, in this case the Weibull distribution, and return the equivalent truncated distribution function as object:

```{r qdist_funs, message=FALSE}
require(qdist)
dtweibull <- dtruncate("weibull") # density function
ptweibull <- ptruncate("weibull") # probability function 
qtweibull <- qtruncate("weibull") # quantile function
rtweibull <- rtruncate("weibull") # random generator function
```

@@

```{r compare_args}
args(dweibull)
args(dtweibull)
```

As `dweibull()` is the function that computes the density function for a Weibull distribution, conventionally named `weibull`, `dtweibull()` is the function that computes the density function for a truncated Weibull distribution, named `tweibull`.

Suppose we want to generate the density function for a **truncated and shifted** Weibull distribution. We need `qshift` package. `qshift` package shares the same logic and the same methodology of `qdist`. The difference is that its functions are designed to generate functions for shifted distributions.
In particular, `qshift` package includes the following four functions:

@@

```{r trunc_shift, message=FALSE}
require(qshift)
dstweibull <- dshift("tweibull")
```

@@

```{r trunc_shift_args}
args(dstweibull)
```

@@

```{r trunc_shift_ex}
x <- seq(0, 5, len = 100)
orig_density <- dweibull(x = x, shape = 1.5, scale = 1)
trunc_density <- dtweibull(x = x, shape = 1.5, scale = 1, L = 1.5, U = 3)
trunc_shift_density <- dstweibull(x = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r trunc_shift_ex_pl, message=FALSE}
df <- data.frame(x=x, 
                 type = factor(c(rep("original density", times=length(x)),
                    rep("trunc density", times=length(x)),
                    rep("trunc-shift density", times=length(x))),
                    levels = c("original density", "trunc density", "trunc-shift density")), 
                  value = c(orig_density, trunc_density, trunc_shift_density))

require(ggplot2)
pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("original density" = "cyan3", 
                                "trunc density" = "#c8a9e3", 
                                "trunc-shift density" = "#47e583")) +
  xlab("Value") + ylab("Density") + ylim(c(0,1.9)) +
  ggtitle(paste("Density plot for original, truncated", 
                "and truncated and shifted Weibull distributions", sep = "\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

@@

Starting from `dshift()` function of `qshift` package, we define `dsweibull()`, the density function of `sweibull`, which is the shifted Weibull distribution:
```{r shift_d}
dsweibull <- dshift("weibull") # density function
```

We have to define also the others functions for shifted distribution, as they are implicitly necessary for further step:

```{r shift_funs}
psweibull <- pshift("weibull") # probability function 
qsweibull <- qshift("weibull") # quantile function
rsweibull <- rshift("weibull") # random generator function
```

The density function of a shifted and truncated Weibull distribution is generated setting `sweibull` distribution as `ptruncate()` argument:
```{r shift_trunc_d}
dtsweibull <- dtruncate("sweibull")
```

@@

```{r shift_trunc_d_ex}
orig_density <- dweibull(x = x, shape = 1.5, scale = 1)
shift_density <- dsweibull(x = x, shape = 1.5, scale = 1, shift = 1)
shift_trunc_density <- dtsweibull(x = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r shift_trunc_ex_pl}
df <- data.frame(x=x, 
                 type = factor(c(rep("original density", times=length(x)),
                    rep("shift density", times=length(x)),
                    rep("shift-trunc density", times=length(x))),
                    levels = c("original density", "shift density", "shift-trunc density")), 
                  value = c(orig_density, shift_density, shift_trunc_density))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("original density" = "cyan3", 
                                "shift density" = "#39375b", 
                                "shift-trunc density" = "#FF89B4")) +
  xlab("Value") + ylab("Density") + ylim(c(0,1.9)) +
  ggtitle(paste("Density plot for original, shifted and",
                "shifted and truncated Normal distributions", sep = "\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

@@

Let us see graphically the differences between the density function of original Weibull distribution, of truncated and shifted Weibull distribution and of shifted and truncated Weibull distribution. We refers to values computed in the previous examples.

```{r ts_st_ex_pl}
df <- data.frame(x=x, 
                 type = factor(c(rep("original density", times=length(x)),
                    rep("trunc-shift density", times=length(x)),
                    rep("shift-trunc density", times=length(x))),
                    levels = c("original density", "trunc-shift density", "shift-trunc density")), 
                  value = c(orig_density, trunc_shift_density, shift_trunc_density))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("original density" = "cyan3", 
                                "trunc-shift density" = "#47e583", 
                                "shift-trunc density" = "#FF89B4")) +
  xlab("Value") + ylab("Density") + ylim(c(0,1.9)) +
  ggtitle(paste("Density plot for original, truncated-shifted", 
                "and shifted-truncated Weibull distribution", sep="\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

@@

Looking at the figure we can see that a truncated and shifted distribution is different from a shifted and truncated one.

Let us analyze the other distribution functions:

@@

```{r shift_trunc_p}
# probability distribution function of a truncated and shifted Weibull distribution
pstweibull <- pshift("tweibull")
# probability distribution function of a shifted and truncated Weibull distribution
ptsweibull <- ptruncate("sweibull")
```

```{r shift_trunc_p_ex}
orig_distr <- pweibull(q = x, shape = 1.5, scale = 1)
trunc_shift_distr <- pstweibull(q = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
shift_trunc_distr <- ptsweibull(q = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r shift_trunc_p_ex_pl}
df <- data.frame(x=x, 
                 type = factor(c(rep("original distr", times=length(x)),
                    rep("shift-trunc distr", times=length(x)),
                    rep("trunc-shift distr", times=length(x))),
                    levels = c("original distr", "shift-trunc distr", "trunc-shift distr")), 
                  value = c(orig_distr, shift_trunc_distr, trunc_shift_distr))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("original distr" = "cyan3", 
                                "shift-trunc distr" = "#FF89B4", 
                                "trunc-shift distr" = "#47e583")) +
  xlab("Quantile") + ylab("Probability") +
  ggtitle(paste("Distribution plot for original, truncated-shifted", 
                "and shifted-truncated Weibull distribution", sep="\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

@@

```{r shift_trunc_q}
# quantile function of a truncated and shifted Weibull distribution
qstweibull <- qshift("tweibull")
# quantile function of a shifted and truncated Weibull distribution
qtsweibull <- qtruncate("sweibull")
```

```{r shift_trunc_q_ex}
x <- 1:999/1000
orig_quant <- qweibull(p = x, shape = 1.5, scale = 1)
trunc_shift_quant <- qstweibull(p = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
shift_trunc_quant <- qtsweibull(p = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r shift_trunc_q_ex_pl}
df <- data.frame(x=x, 
                 type = factor(c(rep("original quant", times=length(x)),
                    rep("shift-trunc quant", times=length(x)),
                    rep("trunc-shift quant", times=length(x))),
                    levels = c("original quant", "shift-trunc quant", "trunc-shift quant")), 
                  value = c(orig_quant, shift_trunc_quant, trunc_shift_quant))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("original quant" = "cyan3", 
                                "shift-trunc quant" = "#FF89B4", 
                                "trunc-shift quant" = "#47e583")) +
  xlab("Probability") + ylab("Quantile") +
  ggtitle(paste("Quantile plot for original, truncated-shifted", 
                "and shifted-truncated Weibull distribution", sep="\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

@@

```{r shift_trunc_r}
# random generation function of a truncated and shifted Weibull distribution
rstweibull <- rshift("tweibull")
# random generation function of a shifted and truncated Weibull distribution
rtsweibull <- rtruncate("sweibull")
```

```{r shift_trunc_r_ex}
orig_rand <- rweibull(n=1000, shape = 1.5, scale = 1)
trunc_shift_rand <- rstweibull(n=1000, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
shift_trunc_rand <- rtsweibull(n=1000, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r shift_trunc_r_ex_pl}
df <- data.frame(x=1:1000, 
                 type = factor(c(rep("original rand", times=1000),
                    rep("shift-trunc rand", times=1000),
                    rep("trunc-shift rand", times=1000)),
                    levels = c("original rand", "shift-trunc rand", "trunc-shift rand")), 
                  value = c(orig_rand, shift_trunc_rand, trunc_shift_rand))

pl <- ggplot(data = df, mapping = aes(x=value, group=type, fill=type)) +
  geom_density(alpha=0.6) +
  scale_fill_manual(values = c("original rand" = "cyan3",
                              "shift-trunc rand" = "#FF89B4",
                              "trunc-shift rand" = "#47e583")) +
  ggtitle(paste("Density plot for original, truncated-shifted", 
                "and shifted-truncated Weibull distribution", sep="\n")) +
  xlab("Quantile") + ylab("Density") +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

@@

```{r l_trunc_shift}
ltweibull <- function(x, shift = 0 ,L = -Inf, U = Inf){
  x1 <- x - shift
  # starting values for parameters (scale, shape)
  shape <- (sd(x1)/mean(x1))^(-1.086)
  scale <- mean(x1)/gamma(1+1/shape)
  # parameters vector definition
  theta <- c(shape, scale)
  # likelihood function
  ll <- function(theta, x, shift = 0, L = -Inf, U = Inf){
    shape <- theta[1]
    scale <- theta[2]
    ld <- dstweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U, log=TRUE)
    -sum(ld)
  }
  # maximum likelihood estimation
  optim(par = theta, fn = ll, x = x, shift=shift, L = L, U = U, method = "Nelder-Mead")[["par"]]
}
```

@@

```{r l_trunc_shift_ex}
x <- rstweibull(n = 100000, shape = 1.6, scale = 1, shift = 1, L = 1, U = 2.5)
# maximum likelihood estimate for the shifted and truncated weibull distribution
ltweibull(x = x, shift=1, L = 1, U = 2.5)
```

@@

We will consider the previously defined density function, `dtsweibull()`, for a shifted and truncated Weibull distribution and we use it inside the estimator function:

```{r l_shift_trunc}
ltweibull <- function(x, shift = 0, L = -Inf, U = Inf){
  x1 <- x - shift
  # starting values for parameters (scale, shape)
  shape <- (sd(x1)/mean(x1))^(-1.086)
  scale <- mean(x1)/gamma(1+1/shape)
  # parameters vector definition
  theta <- c(shape, scale)
  # likelihood function
  ll <- function(theta, x, L = -Inf, U = Inf){
    shape <- theta[1]
    scale <- theta[2]
    ld <- dtsweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U, log=T)
    -sum(ld)
  }
  # maximum likelihood estimation
  optim(par = theta, fn = ll, x = x, L = L, U = U, method = "Nelder-Mead")[["par"]]
}
```

@@

```{r l_shift_trunc_ex}
x <- rtsweibull(n = 100000, shape = 1.6, scale = 1, shift = 1, L = 0.5, U = 3)
# maximum likelihood estimate for the shifted weibull distribution
ltweibull(x = x, shift = 1, L = 0.5, U = 3)
```
